---
title: "Exercise 09 Student Survey"
author: "Taniya Adhikari"
date: "10/3/2020"
output: rmarkdown::github_document
---


Importing the dataset

```{r, echo=FALSE}

library(knitr)
survey_df <- read.csv("data/student-survey.csv")
kable(head(survey_df,5),caption = "Table 9.1: A table of the Dataset with first 5 values.")


```


**a. Use R to calculate the covariance of the Survey variables and provide an explanation of why you would use this calculation and what the results indicate.**  

**Answer a.**  
  
```{r}

kable(cov(survey_df),caption = "Table 9.2: Covariance of variables.")

```
    
Covariance is a good way to show how two variables are related with respect to its mean. Meaning if one variable deviates from its mean, then other variable should also change in similar way, i.e positive or negative. A positive covariance means both variable deviates from the mean in same direction and a negative covariance means that both variable deviates from the mean in opposite direction.  
      
In the above table, a positive covariance between **TimeTV** and **Happiness** indicates that as the Time spent watching TV increases, the Happiness level also increases both deviates in the same direction. A negative covariance between **TimeReading** and **TimeTV** indicates that as the time spent on reading increases, time spent on watching tv decreases both deviates in opposite direction.  
      
**b. Examine the Survey data variables. What measurement is being used for the variables? Explain what effect changing the measurement being used for the variables would have on the covariance calculation. Would this be a problem? Explain and provide a better alternative if needed.**  
     
**Answer b.**  
   
In **Table 9.1**  
1. TimeReading is in hours and continuous variable.  
2. TimeTV is in minutes and continuous variable.  
3. Happiness is an integer and discrete variable with unknown measurment unit.  
4. Gender is a binary variable with 0 or 1 is either male or female.  
  
**TimeReading and TimeTV spent are not in same scale of measurement.**  
  
```{r}

## converting the TimeReading variable 
survey_df$TimeReading_min <- survey_df$TimeReading * 60
kable(head(survey_df,5),caption = "Table 9.3: A table of the Dataset with first 5 values.")

kable(cov(survey_df),caption = "Table 9.4: Covariance of variables with measurement change.")
```
    
The covariance is a good way to check the nature of relationship between the variables, i.e positive or negative. However, covariance is not the best measure for relationship because it changes as the scale of measurement changes. In the above table, I compared TimeReading in hours with TimeTV in minutes and TimeReading in minutes with TimeTV in minutes, the covariance changed from -20.36 to -1221.82, but the nature of relationship didn't change. Therefore Covariance cannot be compared with data in different scales.  
   
Covariance can be used if all the variables were measured in same units, but when variables cannot be converted in same scale (such as Happiness cannot be converted into minutes) covariance cannot be used. A better way is to standardize it where any scale of measurement is converted. A correlation coefficient is a better measurement of relationship between the two variables. 

```{r}

kable(cor(survey_df),caption = "Table 9.5: Correlation coefficient of variables with measurement change.")

```
     
Above table shows the correlation coefficient does not change as the scale of measurement changes.  
    
**c. Choose the type of correlation test to perform, explain why you chose this test, and make a prediction if the test yields a positive or negative correlation?**  
    
**Answer c.**  
   
```{r, warning = FALSE, echo=FALSE}
cor.test(survey_df$TimeReading_min, survey_df$TimeTV)

cor.test(survey_df$TimeReading_min, survey_df$TimeTV, alternative = "less", method = "spearman")

cor.test(survey_df$TimeReading_min, survey_df$TimeTV, alternative = "less", method = "kendall")

```
  
  
To choose the test between parametric and non-parametric, I would check the sample size and normality.  
  
  
```{r, warning = FALSE, echo=FALSE}
library(pastecs)
str(survey_df)
stat.desc(cbind(survey_df$TimeReading_min, survey_df$TimeTV), basic = FALSE, norm = TRUE)

```
  
Although, central limit theorem suggests that data becomes normal as it gets bigger. The Pearson correlation coefficient test is not ideal for such small sample size (n=11), as it's violating the normality assumption (skew and kurtosis is not 0). For Pearson correlation coefficient $r$ to be significant, sample size has to be large enough (n $\geq$ 30). $r$ changes as the sample size gets bigger. Kendall's $\tau$ is probably not the ideal test either, because we don't have large number of tied ranks.  
  
I will use Spearman's $r_{s}$ because it is similar to Pearson coefficient $r$ but works with ranks and works better for smaller sample size. Spearman works for both normal and non-normal distribution. The p-value is also smaller compared to other two tests.

All Tests shows there is  significant strong negative relationship/association between the two variables. Spearman's $r_{s}$ suggests significant monotonic association between the ranks of two variables($r_{s} = -.907$) the Pearson correlation also suggests strong negative relationship ($r = -.88$) and Kendall test ($\tau = -.80$) also suggests strong association of rank.  
  
This means as the Time spent reading increases, the Time spent watching TV decreases. Below graph also suggests the same thing.  

```{r, warning = FALSE, message=FALSE, echo=FALSE}
library(ggplot2)

ggplot(survey_df, aes(x = TimeReading_min, y = TimeTV)) +
  geom_point() + geom_smooth(method=lm, se=FALSE, fullrange=TRUE) + 
  ggtitle("Figure 9.1: Scatter plot of Time Spent Reading vs Watching TV") + 
  xlab("Time Spent Reading (min)") + ylab("Time Spent watching TV (min)")

```
   
**d. Perform a correlation analysis of:**  
**1. All variables**  
**2. A single correlation between two a pair of the variables**  
**3. Repeat your correlation test in step 2 but set the confidence interval at 99%**  
**4. Describe what the calculations in the correlation matrix suggest about the relationship between the variables. Be specific with your explanation.**  
  
**Answer d**    
  
**1.**  
  
```{r, warning = FALSE, message=FALSE, echo=FALSE}

library(corrplot)
corr_R <- cor(survey_df,method = "spearman")
kable(corr_R,caption = "Table 9.6: Spearman Correlation coefficient of variables")

```
    
    
**Figure 9.2: Correlation Plot of Variables**    
  
```{r, warning = FALSE, message=FALSE, echo=FALSE}

corrplot(corr_R, method="circle")

```
  
  
###### For this assignment I will be using Spearman correlation coefficient.  
   
**2.**  
  
**Spearman Correlation Test of Time spent reading vs. Time spent watching TV**
   
```{r, warning = FALSE, message=FALSE, echo=FALSE}

## Correlation analysis between TimeReading_min and TimeTV
cor.test(survey_df$TimeReading_min, survey_df$TimeTV, alternative = "less", method = "spearman")

```
   
**Spearman Correlation Test of Time spent reading vs. Happiness level**  
   
```{r, warning = FALSE, message=FALSE, echo=FALSE}

## Correlation analysis between TimeReading_min and Happiness
cor.test(survey_df$TimeReading_min, survey_df$Happiness, alternative = "less", method = "spearman")

```
  
**Spearman Correlation Test of Time spent watching TV  vs. Happiness level**  
   
```{r, warning = FALSE, message=FALSE, echo=FALSE}

## Correlation analysis between TimeTV and Happiness
cor.test(survey_df$TimeTV, survey_df$Happiness, alternative = "less", method = "spearman")

```
  
**3.**  
  
**Spearman Correlation Test of Time spent reading vs. Time spent watching TV with Confidence level of 99%**
  
```{r, warning = FALSE, message=FALSE, echo=FALSE}

## Correlation analysis between TimeReading_min and TimeTV with 99% Confidence interval

cor.test(survey_df$TimeReading_min, survey_df$TimeTV, alternative = "less", method = "spearman", conf.level = 0.99)

```
  
**Figure 9.3: Confidence level graph for Time spent reading vs. Time spent watching TV**  
  
```{r, warning = FALSE, message=FALSE, echo=FALSE}

library("ggpubr")

## Plots comparing the two confidence level
plot1 <- ggscatter(survey_df, x = "TimeReading_min", y = "TimeTV", add = "reg.line", conf.int = TRUE, cor.coef = TRUE, cor.method = "spearman", xlab = "Time Spent Reading (min)", ylab = "Time Spent watching TV (min)") +ggtitle("Confidence Interval at 95%")

plot2 <- ggscatter(survey_df, x = "TimeReading_min", y = "TimeTV", add = "reg.line", conf.int = TRUE, cor.coef = TRUE, conf.int.level = 0.99, cor.method = "spearman", xlab = "Time Spent Reading (min)", ylab = "Time Spent watching TV (min)")+ggtitle("Confidence Interval at 99%")
ggarrange(plot1, plot2, ncol = 2, nrow = 2)

```

  
**Spearman Correlation Test of Time spent reading vs. Happiness  with Confidence level of 99%**  
   
```{r, warning = FALSE, message=FALSE, echo=FALSE}

## Correlation analysis between TimeReading_min and Happiness with 99% Confidence interval
cor.test(survey_df$TimeReading_min, survey_df$Happiness, alternative = "less", method = "spearman", conf.level = 0.99)

```
  
**Fig 9.4: Confidence level graph for Time spent reading vs. Happiness level**  
  
```{r, warning = FALSE, message=FALSE, echo=FALSE}
## Plots comparing the two confidence level
plot3 <- ggscatter(survey_df, x = "TimeReading_min", y = "Happiness", add = "reg.line", conf.int = TRUE, cor.coef = TRUE, cor.method = "spearman", xlab = "Time Spent Reading (min)", ylab = "Happiness Level") +ggtitle("Confidence Interval at 95%")

plot4 <- ggscatter(survey_df, x = "TimeReading_min", y = "Happiness", add = "reg.line", conf.int = TRUE, cor.coef = TRUE, conf.int.level = 0.99, cor.method = "spearman", xlab = "Time Spent Reading (min)", ylab = "Happiness Level") + ggtitle("Confidence Interval at 99%")
ggarrange(plot3, plot4, ncol = 2, nrow = 2)

```
  
**Spearman Correlation Test of Time spent watching TV vs. Happiness  with Confidence level of 99%**  

```{r, warning = FALSE, message=FALSE, echo=FALSE}

## Correlation analysis between TimeTV and Happiness with 99% Confidence interval
cor.test(survey_df$TimeTV, survey_df$Happiness, alternative = "less", method = "spearman", conf.level = 0.99)

```
    
**Figure 9.5: Confidence level graph for Time spent watching TV vs. Happiness level**  
  
```{r, warning = FALSE, message=FALSE, echo=FALSE}
## Plots comparing the two confidence level
plot5 <- ggscatter(survey_df, x = "TimeTV", y = "Happiness", add = "reg.line", conf.int = TRUE, cor.coef = TRUE, cor.method = "spearman",xlab = "Time Spent watching TV (min)", ylab = "Happiness Level") + ggtitle("Confidence Interval at 95%")

plot6 <- ggscatter(survey_df, x = "TimeTV", y = "Happiness", add = "reg.line", conf.int = TRUE, cor.coef = TRUE, conf.int.level = 0.99, cor.method = "spearman",xlab = "Time Spent watching TV (min)", ylab = "Happiness Level") + ggtitle("Confidence Interval at 99%")
ggarrange(plot5, plot6, ncol = 2, nrow = 2)

```
  
**4.**  
  
Using Table 9.6 there is a significant strong negative relationship between TimeReading and TimeTV with Spearman $r_{s} = -.91$. There is a significant moderate negative relationship between TimeReading and Happiness with Spearman $r_{s} = -.41$.There is a significant positive relationship between TimeTV and Happiness with Spearman $r_{s} = .57$.  
   
This means as students spend more time reading, they will also decrease their time watching TV and their happiness level also decreases. And as students spend more time watching TV they are more likely to be happier.  
   
**e. Calculate the correlation coefficient and the coefficient of determination, describe what you conclude about the results.**  
   
**Answer e**  
  
```{r, warning = FALSE, message=FALSE, echo=FALSE}
## correlation coefficient
r <- cor(survey_df$TimeReading_min, survey_df$TimeTV, method = "spearman")
r

##coefficient of determination
rsquared <- r^2
rsquared

```
    
The spearman correlation coefficient $r_{s} = -0.91$ and correlation of determination $r_{s}^2 = 0.82$. The $r_{s}^2$ is the variation in the ranks of the two variables. This means that Time spent Reading is highly correlated with Time spent watching TV and it can account for 82% of the variation in Time spent watching TV.   
    
  
**f. Based on your analysis can you say that watching more TV caused students to read less? Explain.**  
  
**Answer f**  
  
Yes, The relationship between the two are in negative direction, the relationship is strong because of $r_{s} = -0.91$ and the significance value is $p = 0.00005761$ is less than .05 which means watching more TV causes students to read less.  
   
   
**g. Pick three variables and perform a partial correlation, documenting which variable you are “controlling”. Explain how this changes your interpretation and explanation of the results.**  
   
   
**Answer g**  

In Table 9.6, we know that TimeReading is negatively related to TimeTV and TimeTV is positively related to Happiness, but Happiness is negatively related to TimeReading. So for partial correlation I choose the Happiness as a control variable to measure the pure relationship between TimeReading and TimeTV.  

###### For this part of the question I will be using Pearson correlation coefficient for comparison. I couldn't find a way to do partial correlation in spearman test.  
  

```{r, warning = FALSE, message=FALSE, echo=FALSE}

library(ggm)
library(ppcor)
#pc <- pcor(c("TimeReading_min", "TimeTV", "Happiness"), var(survey_df))
#pc^2
#pc
#pcor.test(pc,1,11)

```

The following code gave me the output: 

> pc <- pcor(c("TimeReading_min", "TimeTV", "Happiness"), var(survey_df))  
> pc^2  
[1] 0.762033  
> pc  
[1] -0.872945  
> pcor.test(pc,1,11)  
  
$tval  
[1] -5.061434  
  
$df  
[1] 8  
  
$pvalue  
[1] 0.0009753126  
  
The $r$ actually reduced from -.88 to -.87 and r-squared got reduced from 0.78 to 0.76. Although, the correlation still seems significant, the correlation between the two variables is decreased. Before the partial correlation, the effects of Happines, the TimeReading and TimeTV spent shared a 78% of the variation. Now it is reduced by 2%. This explains that Time spent Reading alone still shares higher variation in Time spent watching TV

