---
title: "Exercise 13"
author: Taniya Adhikari"
date: 10/25/2020
output: word_document
---

## **Data Set**  
   
   
   
   
```{r, echo = FALSE, warning=FALSE}
library(knitr)
library(foreign)
library(tinytex)
surgery_df <- read.arff("data/ThoraricSurgery.arff")
kable(head(surgery_df,5),caption = "Table 13.1: A table of the Dataset with first 5 values.")

```
    
        
## **Summary Statistics**   
    
       
         
```{r, echo = FALSE, warning=FALSE}
model.1 <- glm(Risk1Yr ~ ., family = binomial, data = surgery_df)
summary(model.1)
```

      
         
           
According to the summary, variable PRE9 - Dyspnoea before surgery being True and PRE14 - size of the original tumour OC14 are highly significant variables with p-value < 0.01 and has greater impact of Risk1Yr being True. PRE17 - Type 2 DM - diabetes mellitus being True and PRE30 - Smoking being True are also significant with p-value < 0.05. All the other variables has p-value between 0.05 < p-value < 1, indicating lesser significance. The value of the deviance for this model (341.19) is also less than the null model (395.61).  
   
       
         
          
          
```{r, echo = FALSE, warning=FALSE}
library(pander)
surgery_df$predicted = predict(model.1, newdata=surgery_df, type="response")
pander(table(surgery_df$Risk1Yr, surgery_df$predicted> 0.5))

```
     
       
         
          
           
It seems like the model is good at predicting false values, with the accuracy rate of 85.3%. However the accuracy rate of True values of predicting the 23.07%. Suggests that model needs improvement, maybe needs dimensionality reduction.  